---
title: "Salary Analysing"
author: "Safa Alshaalan"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
surveyresult <- read.csv("survey_results_public.csv")
surveyquestions <- read.csv("survey_results_schema.csv")
library(plyr) # To use count
library(ggplot2) # Using it with the visualization
library(knitr) # To organize the tables
library(rmarkdown) # To organize the tables
library(dplyr) # To use %>%
library(visdat) # To use vis_miss
library(tidyr) # To use drop_na
library(tidyverse)
library(cowplot)
library(reshape2)
```

## About the project
The main research question in this project is "What is the Predicted salary for the software developers ?". To answer the main question I'll analyse a dataset that I got from Kaggle "<https://www.kaggle.com/stackoverflow/so-survey-2017>"

A short brief about the dataset:
It is a look into the lives of over 64,000 developers from 213 countries.

The data is made up of two files, but I'll be using only the public:

survey_results_public.csv - CSV file with main survey results, one respondent per row and one column per answer.
```{r, echo=FALSE}
paged_table(surveyresult, options = list(rows.print = 15))
```

survey_results_schema.csv - CSV file with survey schema, i.e., the questions that correspond to each column name.
```{r, echo=FALSE}
paged_table(surveyquestions, options = list(rows.print = 15))
```

The majority of the questions are about the developer's career, such as the languages they worked on, what are their professions and what kind of developers they are, if they web developers, applications developers etc.


## EDA and Features engineering:

Check if I have NA's in the dataset
```{r}
sum(is.na(surveyresult))
```

Visualizing the missing values
```{r, echo=FALSE}
vis_miss(surveyresult, warn_large_data = FALSE)
```

Because the dataset includes large missing values, I'll create a dataframe with the columns I'll be using.

Executing the columns and creating the dataframe under name "DFSR"
```{r, include=FALSE}
 DFSR <- surveyresult%>%
  select(Professional,EmploymentStatus,FormalEducation,YearsProgram,YearsCodedJob,UnderstandComputers,JobProfile,HaveWorkedLanguage,HaveWorkedDatabase,HaveWorkedFramework,HaveWorkedPlatform,Salary)
```

```{r, echo=FALSE}
paged_table(DFSR, options = list(rows.print = 15))
```

Chcking the NA's in the dataframe "DFSR"
```{r}
sum(is.na(DFSR))
```

Visualizing the missing values in the dataframe "DFSR"
```{r}
vis_miss(DFSR, warn_large_data = FALSE)
```

The must NA's that I have are in the student's rows because they haven't had a job before, so deleting them will leave me with only **professional developer** category from the **Professional** column.

As we can see in the plot above the most NA's are in the Salary's column, I'll change the NA's with 0
```{r}
DFSR[,c("Salary")][is.na(DFSR[,c("Salary")])] <- 0
```

Replacing the NA's in the **YearsProgram** and **YearsCodedJob** as well with 0, because all of them includes numbers.

```{r}
DFSR[,c("YearsProgram")][is.na(DFSR[,c("YearsProgram")])] <- 0
```

```{r}
DFSR[,c("YearsCodedJob")][is.na(DFSR[,c("YearsCodedJob")])] <- 0
```

Check again the missing values
```{r}
vis_miss(DFSR, warn_large_data = FALSE)
```

Changing the rest of the NA's in the dataframe to None.
```{r}
DFSR[is.na(DFSR)] <- "None"
```

Check if there is any missing values
```{r}
vis_miss(DFSR, warn_large_data = FALSE)
```

After I finished cleaning the dataframe I'll start answering the questions.

-------------------------------------------------------------------------------------

Q1. What are the professions of the software developers in the dataset?

Extracted the **Professional** column.

```{r}
professions <- DFSR[, c("Professional")]
```

Count the categories frequency for the column.
```{r, echo=FALSE}
data.frame(table(unlist(tolower(professions), " ")))
```

Create a dataframe with the result to use it with the visualization.
```{r}
professionsfreq <- data.frame(table(unlist(tolower(professions), " ")))
```

Using the plot to have good visualization results
```{r}
ggplot(data= professionsfreq, aes(x = reorder(Var1, -Freq) ,
                        y = Freq)) +
  geom_col(fill = "#9d8fdb") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  labs(x = "Professions",
       y = "Freq",
       title = "Professions Frequency") +
  theme_minimal()
```
As we can see above the most percentage go to the professional developers. Having a high percentage of that meant that they are the most needed in the market which means that they earn more salary than the others for higher demand for them.

--------------------------------------------------------------------------------------------------

Q2. What are the employment status of the software developers in the dataset?

Extracting the **EmploymentStatus** column as a dataframe to count the frequency
```{r}
EmploymentStatus <- DFSR[, c("EmploymentStatus")]
```

Counting the categories frequency for the employment status and Create a dataframe with the result to use it with the visualization.
```{r}
EmploymentStatusfreq <- data.frame(table(unlist(tolower(EmploymentStatus), " ")))
```

```{r}
EmploymentStatusfreq
```

Using the plot to have good visualization results
```{r}
ggplot(data= EmploymentStatusfreq, aes(x = reorder(Var1, -Freq) ,
                        y = Freq)) +
  geom_col(fill = "#9d8fdb") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  labs(x = "Professions",
       y = "Freq",
       title = "Professions Frequency") +
  theme_minimal()
```
As we can see above we have 7 employee status in the data.

-------------------------------------------------------------------------------------

Q3.	What are the formal educations of the software developers in the dataset?

Extracting the **FormalEducation** column as a dataframe to count the frequency
```{r}
FormalEducation <- DFSR[, c("FormalEducation")]
```

Counting the categories frequency for the formal education and Create a dataframe with the result to use it with the visualization.
```{r}
FormalEducationCategories <- data.frame(table(unlist(tolower(FormalEducation), " ")))
```

```{r}
FormalEducationCategories
```

Using the plot to have good visualization results
```{r}
ggplot(data= FormalEducationCategories, aes(x = reorder(Var1, -Freq) ,
                        y = Freq)) +
  geom_col(fill = "#9d8fdb") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  labs(x = "Professions",
       y = "Freq",
       title = "Professions Frequency") +
  theme_minimal()
```

-----------------------------------------------------------------------------------------

Q4.	What are the most popular languages?

Extracting the **HaveWorkedLanguage** column as a dataframe to count the frequency
```{r}
HaveWorkedLanguage <- DFSR[, c("HaveWorkedLanguage")]
```

The cells includes more than one languages so I'll separate them
```{r, include=FALSE}
lang <- data.frame(do.call('rbind',strsplit(as.character(DFSR$HaveWorkedLanguage),';',fixed = TRUE)))
```

Chick if there is NA's
```{r}
sum(is.na(lang))
```

Remove the NA's
```{r}
lang <- lang[!is.na(lang)]
```

Counting the categories frequency for the languages
```{r}
lang1 <- data.frame(table(unlist(strsplit(tolower(lang), " "))))
```

The first row includes the total number, so I'll delete it as well as the none row
```{r, include=FALSE}
lang2 <- lang1[-c(1,25),]
```

Using the plot to have good visualization results
```{r}
ggplot(data= lang2, aes(x = reorder(Var1, -Freq) ,
                        y = Freq)) +
  geom_col(fill = "#9d8fdb") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  labs(x = "Programming Lnguages",
       y = "Freq",
       title = "Languages Frequency") +
  theme_minimal()
```
As we can see above in the plot we can come up with a conclusion which is the more using languages are the most requested languages in the market these days, which mean if there is a cv with some of the highest languages, that means his salary supposed to be higher than the other.

--------------------------------------------------------------------------------------------------

Q5. What is the average salary for top 5 language?

Following the previous result in the plot, I'll count the average for the salaries for the top 5 languages

1. JavaScript:
Extracting the columns that includes "JavaScript" in HaveWorkedLanguage column. 
```{r}
JavaScript <- paged_table(dplyr::filter(DFSR, grepl("JavaScript",HaveWorkedLanguage)))
```

Extract the Salary column from "JavaScript"
```{r}
JavaScriptAS <- JavaScript[, c("Salary")]
```

Count the average
```{r}
JavaScriptASmean <- print(mean(JavaScriptAS))
```


2. SQL:
Extracting the columns that includes "SQL" in HaveWorkedLanguage column. 
```{r}
SQL <- paged_table(dplyr::filter(DFSR, grepl("SQL",HaveWorkedLanguage)))
```

Extract the Salary column from "SQL"
```{r}
SQLAS <- SQL[, c("Salary")]
```

Count the average
```{r}
SQLASmean <- print(mean(SQLAS))
```


3. Java:
Extracting the columns that includes "Java" in HaveWorkedLanguage column. 
```{r}
Java <- paged_table(dplyr::filter(DFSR, grepl("Java",HaveWorkedLanguage)))
```

Extract the Salary column from "Java"
```{r}
JavaAS <- Java[, c("Salary")]
```

Count the average
```{r}
JavaASmean <- print(mean(JavaAS))
```


4. C#:
Extracting the columns that includes "C#" in HaveWorkedLanguage column. 
```{r}
C <- paged_table(dplyr::filter(DFSR, grepl("C#",HaveWorkedLanguage)))
```

Extract the Salary column from "C#"
```{r}
CAS <- C[, c("Salary")]
```

Count the average
```{r}
CASmean <- print(mean(CAS))
```


5. Python:
Extracting the columns that includes "Python" in HaveWorkedLanguage column. 
```{r}
Python <- paged_table(dplyr::filter(DFSR, grepl("Python",HaveWorkedLanguage)))
```

Extract the Salary column from "Python"
```{r}
PythonAS <- Python[, c("Salary")]
```

Count the average
```{r}
PythonASmean <- print(mean(PythonAS))
```

Compare the salaries mean for each language
```{r}
means <- data.frame(x=c(20048.95,19328.11,19355.74,21483.68,18127.87),
                     y=c("JavaScript","SQL","Java","C#","Python"))
```

```{r}
ggplot(data= means, aes(x = reorder(y, -x) ,
                        y = x)) +
  geom_col(fill = "#9d8fdb") +
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  labs(x = "Lnguages",
       y = "Salary",
       title = "Salaries mean comparetion") +
  theme_minimal()
```
As we can see above the salary average for "C#" language is the highest of the salaries. Which mean hiring a person who knows that language will require the company to give him a higher salary for that.

---------------------------------------------------------------------------------------------

Regression:

```{r}
test1 <- lm(Salary ~ YearsProgram, data = DFSR)
```

```{r}
sigma(test1)
```

```{r}
sigma(test1)^2
```

```{r}
Reg1 <- lm(Salary ~ YearsProgram + YearsCodedJob, data = DFSR)
```

```{r}
Reg2 <- lm(Salary ~ ., data = DFSR)
```

```{r}
paged_table(broom::tidy(Reg2))
```

```{r}
p1 <- ggplot(DFSR, aes(YearsProgram, Salary)) + 
  geom_point(size = 1, alpha = .4) +
  geom_smooth(se = FALSE) +
  scale_y_continuous("Salary") +
  ggtitle(paste("Non-transformed variables with a\n",
                "non-linear relationship."))
```

```{r}
p2 <- ggplot(DFSR, aes(YearsProgram, Salary)) + 
  geom_point(size = 1, alpha = .4) + 
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_log10("Salary") +
  xlab("YearsProgram") +
  ggtitle(paste("Transforming variables can provide a\n",
                "near-linear relationship."))
```

```{r}
gridExtra::grid.arrange(p1, p2, nrow = 1)
```




























